Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2021-10-15 doc/AMS-15/AMS-15_poster.Rmd
Add text. Revert to hgis - graphite for Haiti comparison plot. Fix colors. Add core top to BChron.

1. Acidify
  - ~30mg Carbonate 
  - produces 6mL CO<sub>2</sub> @ ~1ATM
1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Standard: live gas in vials
- Blank: C-1 or TIRI-F in vials

@@

@@

1. Correct for fractionation using online <sup>13</sup>C.
1. Determine per-sample mean and error.
1. Normalize to mean of Gas standard.
1. Blank correct: LBC using mean of carbonate blanks

@@

@@

# Current independent blank

@@

#### Linear fit used to correct:

@@

@@

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  ungroup()

@@

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr)
```

@@

```{r, eval = FALSE}
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

```{r}
# No outliers
combined_results %>% 
  filter(abs(fm_diff) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Core depth (cm)",
       y = "Fm - mean Fm")
```

@@

cal_sub <- combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method)) %>% 
  select(id, depth, rc_age, sig_rc_age, method)

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gr <- cal_sub %>% 
  filter(method == "graphite") %>% 
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gas <- cal_sub %>% 
  filter(method == "hgis") %>%
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

@@

```{r}
plot(cal_dates)
```

>>>>

### Sample Preparation

1. 10-30mg Carbonate in septa-sealed 7mL vials
  - produces 2-6mL CO<sub>2</sub> @ ~1ATM
1. Evacuate and Acidify with 1 mL H3PO4 using automated Gilson prep bench
1. Add helium to 1 ATM

### Measurement

1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Normalizing standards: live tank gas in vials
- Blanks: C-1 or TIRI-F in vials

@@

![HGIS Photo](images/gas_control.jpg)
![HGIS Photo](images/gilson.png)

@@

1. Correct for fractionation using online AMS <sup>13</sup>C measurement.
1. Determine per-sample mean and error (repeatabilty and counting error).
1. Normalize to mean of Gas standard.
1. Blank correct: Blank subtraction using mean of carbonate blanks

@@

Introducing sample through an open split using helium displacement results in a decreasing flow of CO2 to the source over time. Measured ratio can depend on sample current, so CO2 flow rate must be kept in a range where ion currents are relatively constant with changing flow. We evaluated C- current over a range of CO2 flows and found that a range of flows produce a "flat top" in current. Less CO2 will starve the source of available material to dissociate and ionize, and too much CO2 also suppresses ionization.

@@

#### Current dependent blank

Measurements of live and dead gas over a range of C- currents show that there is a current-dependent component to the blank. We believe this is due to a combination of a constant amount of contaminant CO2 and a "target blank" current. This current dependent blank follows the constant contamination model and can be represented with a current balance between the sample and the constant blank.

#### Current independent blank

@@

While this current-dependent blank is important to consider for samples producing varying ion current, we simplified our assessment of blank by keeping sample currents as constant as possible for blanks, normalizing standards, and unknowns, which allows a simpler correction depending only on the Fm of the blank, standard, and unknown.

This leads to the following model for blank correction:

@@

Data from the measurement of consensus standards over a range of fraction modern shows that this simple correction is effective for removing blank contribution without bias.

@@

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff_mean = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  left_join(select(graphite_results, rec_num, fm_corr_gr = fm_corr)) %>% 
  mutate(fm_diff = fm_corr - fm_corr_gr) %>% 
  ungroup()

@@

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr, fm_diff)
```

@@

```{r, eval = FALSE}
ggplot(combined_results) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm")
```

```{r}
# No outliers, vs depth
combined_results %>% 
  filter(abs(fm_diff) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Core depth (cm)",
       y = "Fm - mean Fm")
```

```{r}
# No outliers, vs mean_fm
combined_results %>% 
  filter(abs(fm_diff) < 0.02) %>% 
ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  position = "jitter") +
  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Mean Fm",
       y = "Fm - mean Fm")
```

@@

cal_sub <- combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method),
         cal_curve = "intcal20") %>% 
  select(id, depth, rc_age, sig_rc_age, cal_curve, method) %>%
  add_row(id = "top-1", depth = 0, rc_age = 0, sig_rc_age = 1, cal_curve = "normal", method = "top")

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))

cal_dates_gr <- cal_sub %>% 
  filter(method != "hgis") %>% 
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))

cal_dates_gas <- cal_sub %>% 
  filter(method != "graphite") %>%
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))

@@

```{r, eval=FALSE}
plot(cal_dates)
```
